---
title: "Final Code Workflow for Buffer Strategy 1"
date: "June 16, 2026"
date-format: long
format: html
editor: visual
---

## Loadings

```{r libs, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(pROC)        
library(ROCR)
library(spdep)
library(corrplot)
library(car)
library(grpreg)     
library(dplyr)
library(caret)
library(nortest)
library(DescTools)
library(PRROC)
library(caret)
#Libraries for mapping
library(sf)
library(raster)
library(glmnet)
library(sp)
library(classInt)
library(gstat)
```

## 1A Data Preparation

```{r Data Strategy 1}
data <- read.csv("Buffer-Strategy-1.csv")
```

```{r Standardization}
target_predictors <- c("Elev", "Slope", "PGAComp", "HorizontalCurvature", "VerticalCurvature")
data[, target_predictors] <- scale(data[, target_predictors])
data <- data |> mutate(
  Aspect = as.factor(Aspect),
  Lithology = as.factor(Lithology),
  SoilType = as.factor(SoilType)
)
```




## 1B Diagnostic Checking

```{r Normality test}
# Shapiro-Wilk test
shapiro.test(data$Elev)
shapiro.test(data$Slope)
shapiro.test(data$HorizontalCurvature)
shapiro.test(data$VerticalCurvature)
shapiro.test(data$PGAComp)

# Anderson-Darling test
ad.test(data$Elev)
ad.test(data$Slope)
ad.test(data$HorizontalCurvature)
ad.test(data$VerticalCurvature)
ad.test(data$PGAComp)
```

```{r Correlation Categorical Vars, warning= FALSE}
chisq.test(table(data$Aspect, data$Lithology))
CramerV(table(data$Aspect, data$Lithology))
chisq.test(table(data$Aspect, data$SoilType))
CramerV(table(data$Aspect, data$SoilType))
chisq.test(table(data$SoilType, data$Lithology))
CramerV(table(data$SoilType, data$Lithology))
```

```{r Correlation Numerical Vars}
cor_matrix <- cor(data[, 
                       c("Slope","Elev","HorizontalCurvature","VerticalCurvature","PGAComp")],
                  method = "spearman")

corrplot(cor_matrix, method="number", tl.cex=0.8)
```


## 1C Model Building

### 1C.0

```{r Data Split}
set.seed(123) 
train_index <- createDataPartition(data$Landslide, p = 0.8, list = FALSE)
train_data <- data[train_index, ]
test_data <- data[-train_index, ]
```

```{r Compute for Reference Level}
# Function to compute LS rate per category
ls_rate <- function(var, data) {
  aggregate(data$Landslide, 
            by = list(Category = data[[var]]), 
            FUN = function(x) sum(x) / length(x))
}

# Aspect
aspect_rates <- ls_rate("Aspect", train_data)
colnames(aspect_rates)[2] <- "LS_rate"
print(aspect_rates[order(aspect_rates$LS_rate), ])

# Lithology
lithology_rates <- ls_rate("Lithology", train_data)
colnames(lithology_rates)[2] <- "LS_rate"
print(lithology_rates[order(lithology_rates$LS_rate), ])

# SoilType
soiltype_rates <- ls_rate("SoilType", train_data)
colnames(soiltype_rates)[2] <- "LS_rate"
print(soiltype_rates[order(soiltype_rates$LS_rate), ])
```

```{r Reference Level}
train_data$Aspect    <- relevel(train_data$Aspect, ref = "North")
train_data$Lithology <- relevel(train_data$Lithology, ref = "Recent deposits")
train_data$SoilType  <- relevel(train_data$SoilType, ref = "Loam")    
```

```{r X_train y_train}
X_train <- model.matrix(Landslide ~ Elev + Slope + PGAComp + 
                     VerticalCurvature + HorizontalCurvature + 
                     Aspect + Lithology + SoilType, 
                     data = train_data)[, -1]  
y_train <- train_data$Landslide
```

```{r X_test y_test}
X_test  <- model.matrix(Landslide ~ Elev + Slope + PGAComp + 
                     VerticalCurvature + HorizontalCurvature + 
                     Aspect + Lithology + SoilType, data = test_data)[, -1]
y_test  <- test_data$Landslide
```

```{r group}
cols <- colnames(X_train)

group <- integer(length(cols))

group[grep("^Elev$", cols)]                <- 1
group[grep("^Slope$", cols)]               <- 2
group[grep("^PGAComp$", cols)]             <- 3
group[grep("^HorizontalCurvature$", cols)] <- 4
group[grep("^VerticalCurvature$", cols)]   <- 5
group[grep("^Aspect", cols)]     <- 6
group[grep("^Lithology", cols)]  <- 7
group[grep("^SoilType", cols)]   <- 8

table(group)
cbind(cols, group)
```

### 1C.1 Benchmark Model

```{r standard lr}
benchmark_model <- glm(Landslide ~ Elev + Slope + PGAComp + 
                     VerticalCurvature + HorizontalCurvature + 
                     Aspect + Lithology + SoilType, 
                     data = train_data, family = "binomial")
```

### 1C.2 Proposed Model: grLASSO

```{r grLASSO lr}
# Fit Group LASSO logistic regression
fit_grLASSO <- grpreg(
  X_train, y_train,
  group = group,
  penalty = "grLasso",
  family = "binomial"
)

# Cross-validation to select λ
set.seed(123)
cvfit_grLASSO <- cv.grpreg(
  X_train, y_train,
  group = group,
  penalty = "grLasso",  
  family = "binomial",
  seed = 123
)

lambda_min_grLASSO <- cvfit_grLASSO$lambda.min
cvfit_grLASSO |> plot()
lambda_min_grLASSO

coef_min_grLASSO <- coef(cvfit_grLASSO, lambda = lambda_min_grLASSO)
coef_min_grLASSO
```

### 1C.3 Proposed Model: grSCAD 

```{r grSCAD lr}
# Fit Group LASSO logistic regression
fit_scad <- grpreg(
  X_train, y_train,
  group = group,
  penalty = "grSCAD",
  family = "binomial"
)

# Cross-validation to select λ
set.seed(123)
cvfit_scad <- cv.grpreg(
  X_train, y_train,
  group = group,
  penalty = "grSCAD",   
  family = "binomial",
  seed = 123
)

lambda_min_scad <- cvfit_scad$lambda.min

# Extract coefficients at optimal λ
coef_min_scad <- coef(cvfit_scad, lambda = lambda_min_scad)
coef_min_scad
```




## 1D Model Fitting

### 1D.1 Bootstrapped Benchmark Model

```{r boot standard lr}
B <- 500
coef_mat <- matrix(NA, nrow = B, 
                   ncol = length(coef(
                     glm(Landslide ~ Elev + Slope + PGAComp + VerticalCurvature +
                           HorizontalCurvature + Aspect + Lithology + SoilType, 
                         data = train_data, family="binomial"))))

colnames(coef_mat) <- names(coef(
  glm(Landslide ~ Elev + Slope + PGAComp + VerticalCurvature + 
        HorizontalCurvature + Aspect + Lithology + SoilType, 
      data = train_data, family="binomial")))

set.seed(123)
for (b in 1:B) {
  idx <- sample(seq_len(nrow(train_data)), replace = TRUE)
  d_b <- train_data[idx, ]
  fit_b <- glm(Landslide ~ Elev + Slope + PGAComp + VerticalCurvature + 
                 HorizontalCurvature + Aspect + Lithology + SoilType, 
               data = d_b, family = "binomial")
  coef_b <- coef(fit_b)
  coef_mat[b, names(coef_b)] <- coef_b
}
```

### 1D.2 Bootstrapped Proposed Model: grLASSO 

```{r boot grLASSO lr, eval=FALSE, include=FALSE}
B <- 500   

coef_storage    <- matrix(NA, nrow = ncol(X_train), ncol = B)
colnames(coef_storage) <- paste0("Rep", 1:B)
rownames(coef_storage) <- colnames(X_train)

prediction_list <- vector("list", B)
actual_list     <- vector("list", B)
perf_metrics    <- data.frame(
  Rep = 1:B,
  AUC = NA,
  Accuracy = NA,
  LogLoss = NA,
  Brier = NA
)

set.seed(123)
for (b in 1:B) {
  idx <- sample(seq_len(nrow(X_train)), replace = TRUE)
  Xb <- X_train[idx, ]
  yb <- y_train[idx]
  
  cvfit_b <- cv.grpreg(
    Xb, yb,
    group = group,
    penalty = "grLasso",
    family = "binomial",
    measure = "deviance",
    seed = b
  )
  
  lambda_min_b <- cvfit_b$lambda.min
  
  coef_b <- coef(cvfit_b, lambda = lambda_min_b)[-1]  
  coef_storage[, b] <- coef_b
  
  prob_b <- predict(cvfit_b, Xb, type = "response", lambda = lambda_min_b)
  prediction_list[[b]] <- prob_b
  actual_list[[b]]     <- yb
  
  auc_b <- pROC::auc(pROC::roc(yb, prob_b))
  acc_b <- mean((prob_b > 0.5) == yb)
  logloss_b <- -mean(yb*log(prob_b) + (1-yb)*log(1-prob_b))
  brier_b <- mean((prob_b - yb)^2)
  
  perf_metrics[b, c("AUC","Accuracy","LogLoss","Brier")] <- c(auc_b, acc_b, logloss_b, brier_b)
}

# Save everything together
saveRDS(list(
  coefficients = coef_storage,
  performance  = perf_metrics, 
  actuals      = actual_list
), file = "GroupLASSO-boot-results-strat1.rds")
```

```{r boot grLASSO lr LOAD}

```


### 1D.3 Bootstrapped Proposed Model: grSCAD (copied chunk from 24AprGroupedLASSO-SCAD.qmd)

```{r boot grSCAD lr, eval=FALSE}
B <- 500

coef_storage    <- matrix(NA, nrow = ncol(X_train), ncol = B)
colnames(coef_storage) <- paste0("Rep", 1:B)
rownames(coef_storage) <- colnames(X_train)

prediction_list <- vector("list", B)
actual_list     <- vector("list", B)
perf_metrics    <- data.frame(
  Rep = 1:B,
  AUC = NA,
  Accuracy = NA,
  LogLoss = NA,
  Brier = NA
)

set.seed(123)
for (b in 1:B) {
  idx <- sample(seq_len(nrow(X_train)), replace = TRUE)
  Xb <- X_train[idx, ]
  yb <- y_train[idx]
  
  cvfit_b <- cv.grpreg(
    Xb, yb,
    group   = group,
    penalty = "grSCAD",   
    family  = "binomial",
    measure = "deviance",
    seed    = b
  )
  
  lambda_min_b <- cvfit_b$lambda.min
  
  coef_b <- coef(cvfit_b, lambda = lambda_min_b)[-1]
  coef_storage[, b] <- coef_b
  
  prob_b <- predict(cvfit_b, Xb, type = "response", lambda = lambda_min_b)
  prediction_list[[b]] <- prob_b
  actual_list[[b]]     <- yb
  
  auc_b <- pROC::auc(pROC::roc(yb, prob_b))
  acc_b <- mean((prob_b > 0.5) == yb)
  logloss_b <- -mean(yb*log(prob_b) + (1-yb)*log(1-prob_b))
  brier_b   <- mean((prob_b - yb)^2)
  
  perf_metrics[b, c("AUC","Accuracy","LogLoss","Brier")] <- c(auc_b, acc_b, logloss_b, brier_b)
}

saveRDS(list(
  coefficients = coef_storage,
  performance  = perf_metrics,
  predictions  = prediction_list,
  actuals      = actual_list
), file = "GroupSCAD-boot-results-strat1.rds")
```

```{r boot grSCAD lr LOAD}

```

## 1E Model Diagnostic (incl. Significant Testing)

### 1E.1 Benchmark Model

```{r Multicollinearity}
lm_model <- glm(Landslide ~ Elev + Slope + HorizontalCurvature + VerticalCurvature + 
                     PGAComp + Aspect + Lithology + SoilType, 
                   data = data, family = "binomial"(link = "logit"))


vif(lm_model)
```

```{r boot lr SIGNIFICANT}
ci_lower <- apply(coef_mat, 2, function(x) quantile(x, 0.025, na.rm = TRUE))
ci_upper <- apply(coef_mat, 2, function(x) quantile(x, 0.975, na.rm = TRUE))

coef_median <- apply(coef_mat, 2, function(x) median(x, na.rm = TRUE))
coef_mean   <- apply(coef_mat, 2, function(x) mean(x, na.rm = TRUE))

prop_exclude_zero <- apply(coef_mat, 2, function(x) {
  mean(sign(x) == sign(mean(x, na.rm = TRUE)), na.rm = TRUE)
})

coef_summary <- data.frame(
  term    = colnames(coef_mat),
  mean    = coef_mean,
  median  = coef_median,
  ci_lower = ci_lower,
  ci_upper = ci_upper,
  stringsAsFactors = FALSE
)

coef_summary$mean <- ifelse(prop_exclude_zero >= 0.95,
                           paste0(round(coef_summary$mean, 6), "*"),
                           round(coef_summary$mean, 6))

coef_summary$mean <- ifelse(coef_summary$ci_lower > 0 | coef_summary$ci_upper < 0,
                            paste0(round(coef_summary$mean, 6), "*"),
                            round(coef_summary$mean, 6))


coef_summary <- coef_summary[order(-abs(coef_summary$median)), ]
coef_summary

saveRDS(coef_summary, file = "Boot-Standard-LR-coef-summary-strat1.rds")
```

### 1E.2 Proposed Model: grLASSO

```{r boot grLASSO, eval=FALSE, include=FALSE}
B <- 500   

coef_storage    <- matrix(NA, nrow = ncol(X_train), ncol = B)
colnames(coef_storage) <- paste0("Rep", 1:B)
rownames(coef_storage) <- colnames(X_train)

prediction_list <- vector("list", B)
actual_list     <- vector("list", B)
perf_metrics    <- data.frame(
  Rep = 1:B,
  AUC = NA,
  Accuracy = NA,
  LogLoss = NA,
  Brier = NA
)

set.seed(123)
for (b in 1:B) {
  idx <- sample(seq_len(nrow(X_train)), replace = TRUE)
  Xb <- X_train[idx, ]
  yb <- y_train[idx]
  
  cvfit_b <- cv.grpreg(
    Xb, yb,
    group = group,
    penalty = "grLasso",
    family = "binomial",
    measure = "deviance",
    seed = b
  )
  
  lambda_min_b <- cvfit_b$lambda.min
  
  coef_b <- coef(cvfit_b, lambda = lambda_min_b)[-1] 
  coef_storage[, b] <- coef_b
  
  prob_b <- predict(cvfit_b, Xb, type = "response", lambda = lambda_min_b)
  prediction_list[[b]] <- prob_b
  actual_list[[b]]     <- yb
  
  auc_b <- pROC::auc(pROC::roc(yb, prob_b))
  acc_b <- mean((prob_b > 0.5) == yb)
  logloss_b <- -mean(yb*log(prob_b) + (1-yb)*log(1-prob_b))
  brier_b <- mean((prob_b - yb)^2)
  
  perf_metrics[b, c("AUC","Accuracy","LogLoss","Brier")] <- c(auc_b, acc_b, logloss_b, brier_b)
}

saveRDS(list(
  coefficients = coef_storage,
  performance  = perf_metrics, 
  actuals      = actual_list,
  prediction = prediction_list,
), file = "boot-grLASSO-results-strat1.rds")
```

```{r boot grLASSO READ}
results <- readRDS("boot-grLASSO-results-strat1.rds")

coef_mat <- results$coefficients

coef_mean   <- rowMeans(coef_mat, na.rm = TRUE)
coef_median <- apply(coef_mat, 1, function(x) median(x[x != 0], na.rm = TRUE))
ci_lower    <- apply(coef_mat, 1, function(x) if (any(x != 0)) quantile(x[x != 0], 0.025) else 0)
ci_upper    <- apply(coef_mat, 1, function(x) if (any(x != 0)) quantile(x[x != 0], 0.975) else 0)
inclusion_freq <- rowMeans(coef_mat != 0) * 100

coef_summary_lasso <- data.frame(
  term = rownames(coef_mat),
  mean = coef_mean,
  median = coef_median,
  ci_lower = ci_lower,
  ci_upper = ci_upper,
  inclusion_freq = inclusion_freq,
  stringsAsFactors = FALSE
)

sig_flag <- with(coef_summary_lasso, (ci_lower > 0 | ci_upper < 0))

coef_summary_lasso$mean <- ifelse(sig_flag,
                                    paste0(round(coef_summary_lasso$mean, 3), "*"),
                                    as.character(round(coef_summary_lasso$mean, 3)))

coef_summary_lasso <- coef_summary_lasso[order(-abs(coef_summary_lasso$median)), ]

coef_summary_lasso
```


### 2E.3 Proposed Model: grSCAD

```{r grSCAD}
B <- 500  

coef_storage    <- matrix(NA, nrow = ncol(X_train), ncol = B)
colnames(coef_storage) <- paste0("Rep", 1:B)
rownames(coef_storage) <- colnames(X_train)

prediction_list <- vector("list", B)
actual_list     <- vector("list", B)
perf_metrics    <- data.frame(
  Rep = 1:B,
  AUC = NA,
  Accuracy = NA,
  LogLoss = NA,
  Brier = NA
)


set.seed(123)
for (b in 1:B) {
  idx <- sample(seq_len(nrow(X_train)), replace = TRUE)
  Xb <- X_train[idx, ]
  yb <- y_train[idx]
  
  cvfit_b <- cv.grpreg(
    Xb, yb,
    group   = group,
    penalty = "grSCAD",   
    family  = "binomial",
    measure = "deviance",
    seed    = b
  )
  
  lambda_min_b <- cvfit_b$lambda.min
  
  coef_b <- coef(cvfit_b, lambda = lambda_min_b)[-1]
  coef_storage[, b] <- coef_b
  
  prob_b <- predict(cvfit_b, Xb, type = "response", lambda = lambda_min_b)
  prediction_list[[b]] <- prob_b
  actual_list[[b]]     <- yb
  
  auc_b <- pROC::auc(pROC::roc(yb, prob_b))
  acc_b <- mean((prob_b > 0.5) == yb)
  logloss_b <- -mean(yb*log(prob_b) + (1-yb)*log(1-prob_b))
  brier_b   <- mean((prob_b - yb)^2)
  
  perf_metrics[b, c("AUC","Accuracy","LogLoss","Brier")] <- c(auc_b, acc_b, logloss_b, brier_b)
}

saveRDS(list(
  coefficients = coef_storage,
  performance  = perf_metrics,
  predictions  = prediction_list,
  actuals      = actual_list
), file = "boot-grSCAD-results-strat1.rds")
```


```{r boot grSCAD READ}
results <- readRDS("boot-grSCAD-results-strat1.rds")

coef_storage    <- results$coefficients
prediction_list <- results$predictions
actual_list     <- results$actuals

inclusion_freq <- rowMeans(coef_storage != 0) * 100
sort(inclusion_freq, decreasing = TRUE)

coef_mean_scad   <- rowMeans(coef_storage, na.rm = TRUE)
coef_median_scad <- apply(coef_storage, 1, function(x) median(x[x != 0], na.rm = TRUE))
ci_lower_scad    <- apply(coef_storage, 1, function(x) quantile(x[x != 0], 0.025, na.rm = TRUE))
ci_upper_scad    <- apply(coef_storage, 1, function(x) quantile(x[x != 0], 0.975, na.rm = TRUE))

coef_summary_scad <- data.frame(
  term          = rownames(coef_storage),
  mean          = coef_mean_scad,
  median        = coef_median_scad,
  ci_lower      = ci_lower_scad,
  ci_upper      = ci_upper_scad,
  inclusion_freq= inclusion_freq,
  significant   = sig_flag,
  stringsAsFactors = FALSE
)

sig_flag_scad <- with(coef_summary_scad, (ci_lower > 0 | ci_upper < 0))

coef_summary_scad$mean <- ifelse(sig_flag_scad,
                                         paste0(round(coef_summary_scad$mean, 3), "*"),
                                         as.character(round(coef_summary_scad$mean, 3)))

coef_summary_scad <- coef_summary_scad[order(-abs(coef_summary_scad$median)), ]

coef_summary_scad
```

## 1F Variable Importance

### 1F.1 Benchmark Model

```{r reduced LR model}
benchmark_model_reduced <- glm(Landslide ~  Lithology + SoilType + Slope + PGAComp + Elev, 
                     data = train_data, family = "binomial")
```

```{r LRT on lr}
drop_res <- drop1(benchmark_model_reduced, test="Chisq")

dev_increase <- drop_res$Deviance
names(dev_increase) <- rownames(drop_res)

ranked <- sort(dev_increase, decreasing=TRUE)

total_dev <- sum(ranked, na.rm = TRUE)
percent_importance <- (ranked / total_dev) * 100

rank_table <- data.frame(
  Variable = names(ranked),
  DevianceIncrease = ranked,
  PercentImportance = round(percent_importance, 2),
  Rank = seq_along(ranked)
)
print(rank_table)

drop_res |> print()
```

### 1F.2 Proposed Model: grLASSO

```{r}
coef_summary_lasso$inclusion_freq
```

### 1F.3 Proposed Model: grSCAD

```{r}
coef_summary_scad$inclusion_freq
```




## 1G Model Evaluation

### 1G.1 Benchmark Model

```{r boot lr threshold CODE, eval=FALSE, include=FALSE}
thresholds <- data.frame(Rep = 1:B, F1 = NA_real_)

safe_f1 <- function(preds, y) {
  cm <- table(factor(preds, levels = c(0,1)),
              factor(y, levels = c(0,1)))
  TP <- cm["1","1"]; FP <- cm["1","0"]; FN <- cm["0","1"]
  prec <- ifelse(TP + FP == 0, 0, TP / (TP + FP))
  rec  <- ifelse(TP + FN == 0, 0, TP / (TP + FN))
  if (prec + rec == 0) return(0)
  2 * prec * rec / (prec + rec)
}

for (b in 1:B) {
  coef_b <- coef_mat[b, ]
  names(coef_b) <- colnames(coef_mat)

  intercept <- if ("(Intercept)" %in% names(coef_b)) coef_b["(Intercept)"] else 0
  betas     <- coef_b[names(coef_b) != "(Intercept)"]

  common_vars <- intersect(names(betas), colnames(X_test))
  betas <- betas[common_vars]
  X_test_num <- as.matrix(X_test[, common_vars, drop = FALSE])

  lp <- drop(X_test_num %*% betas + intercept)
  probs <- 1 / (1 + exp(-lp))

  thresh_seq <- seq(0.01, 0.99, by = 0.01)
  f1_scores <- sapply(thresh_seq, function(t) {
    preds_t <- ifelse(probs > t, 1, 0)
    safe_f1(preds_t, y_test)
  })

  thresholds$F1[b] <- thresh_seq[which.max(f1_scores)]
}

saveRDS(thresholds, "boot-lr-threshold-strat1.rds")
```

```{r boot lr threshold READ}
thresholds <- readRDS("boot-lr-threshold-strat1.rds")

summary(thresholds$F1)
median_thresh_lr <- median(thresholds$F1, na.rm = TRUE)
```

```{r LR evaluation metrics}
evaluate_model <- function(model, data, threshold = median_thresh_lr) {
  probs <- predict(model, newdata = test_data, type = "response")
  preds <- ifelse(probs > threshold, 1, 0)
  
  cm <- table(Predicted = preds, Actual = data$Landslide)
  
  TP <- cm[2,2]; TN <- cm[1,1]
  FP <- cm[2,1]; FN <- cm[1,2]
  
  accuracy <- (TP + TN) / sum(cm)
  sensitivity <- TP / (TP + FN)
  specificity <- TN / (TN + FP)
  precision <- TP / (TP + FP)
  f1 <- 2 * precision * sensitivity / (precision + sensitivity)
  
  brier <- mean((probs - data$Landslide)^2)
  logloss <- -mean(data$Landslide * log(probs) + 
                   (1 - data$Landslide) * log(1 - probs))
  
  # Build ROC object
  roc_obj <- pROC::roc(data$Landslide, probs)
  auc_val <- pROC::auc(roc_obj)  
  
  pr_obj <- PRROC::pr.curve(
  scores.class0 = probs[data$Landslide == 1],
  scores.class1 = probs[data$Landslide == 0],
  curve = FALSE
  )
  pr_auc <- pr_obj$auc.integral

  
  best_thresh <- coords(roc_obj, "best", best.method = "youden", ret = "threshold")
  f1_thresh <- coords(roc_obj, "best", best.method = "closest.topleft", ret = "threshold")

  
  return(list(Accuracy = accuracy,
              Sensitivity = sensitivity,
              Specificity = specificity,
              Precision = precision,
              F1 = f1,
              PRAUC = pr_auc,
              AUC = auc_val,
              Brier = brier,
              LogLoss = logloss
              ))
}

metricsLR <- evaluate_model(benchmark_model_reduced, test_data)
metricsLR
```


### 1G.2 Proposed Model: grLASSO

```{r reduced grLASSO model}
X_train_grLASSO_reduced <- model.matrix(Landslide ~ Elev + Slope + PGAComp +  Lithology + 
                                          Aspect + SoilType,
                     data = train_data)[, -1]  
X_test_grLASSO_reduced <- model.matrix(Landslide ~ Elev + Slope + PGAComp +  Lithology + 
                                          Aspect + SoilType,
                                       data = test_data)[, -1]

cols <- colnames(X_train_grLASSO_reduced)

group_grLASSO_reduced <- integer(length(cols))

group_grLASSO_reduced[grep("^Slope$", cols)]    <- 1
group_grLASSO_reduced[grep("^PGAComp$", cols)]  <- 2
group_grLASSO_reduced[grep("^Elev", cols)] <- 3
group_grLASSO_reduced[grep("^Lithology", cols)] <- 4
group_grLASSO_reduced[grep("^Aspect", cols)] <- 5
group_grLASSO_reduced[grep("^SoilType", cols)] <- 6

table(group_grLASSO_reduced)
cbind(cols, group_grLASSO_reduced)


fit_grLASSO_reduced <- grpreg(
  X_train_grLASSO_reduced, y_train,
  group = group_grLASSO_reduced,
  penalty = "grLasso",
  family = "binomial"
)

set.seed(123)
cvfit_grLASSO_reduced <- cv.grpreg(
  X_train_grLASSO_reduced, y_train,
  group = group_grLASSO_reduced,
  penalty = "grLasso",  
  family = "binomial",
  seed = 123
)

lambda_min_grLASSO_reduced <- cvfit_grLASSO_reduced$lambda.min
cvfit_grLASSO_reduced |> plot()
lambda_min_grLASSO_reduced


coef_min_grLASSO_reduced <- coef(cvfit_grLASSO_reduced, lambda = lambda_min_grLASSO_reduced)
coef_min_grLASSO_reduced    
```

```{r  boot grLASSO threshold CODE}
results <- readRDS("GroupLASSO-boot-results-strat1.rds")

coef_storage    <- results$coefficients  
actual_list     <- results$actuals       

inclusion_freq <- rowMeans(coef_storage != 0) * 100

selected_levels <- rownames(coef_storage)[inclusion_freq >= 80]

X_train_reduced <- as.matrix(X_train[, intersect(colnames(X_train), selected_levels), drop = FALSE])

cv_lasso_reduced <- cv.glmnet(
  X_train_reduced, y_train,
  family = "binomial", alpha = 1,
  standardize = FALSE
)

X_test <- as.data.frame(X_test)
missing_cols <- setdiff(colnames(X_train_reduced), colnames(X_test))
for (mc in missing_cols) {
  X_test[[mc]] <- 0
}
X_test_reduced <- as.matrix(X_test[, colnames(X_train_reduced), drop = FALSE])

stopifnot(nrow(X_test_reduced) == length(y_test))


thresholds <- data.frame(Rep = 1:ncol(coef_storage), F1 = NA_real_)

safe_f1 <- function(preds, y) {
  cm <- table(
    Predicted = factor(preds, levels = c(0,1)),
    Actual    = factor(y, levels = c(0,1))
  )
  TP <- cm["1","1"]; FP <- cm["1","0"]; FN <- cm["0","1"]
  prec <- ifelse(TP + FP == 0, 0, TP / (TP + FP))
  rec  <- ifelse(TP + FN == 0, 0, TP / (TP + FN))
  if (prec + rec == 0) return(0)
  2 * prec * rec / (prec + rec)
}

for (b in 1:ncol(coef_storage)) {
  coef_b <- coef_storage[, b]

  intercept <- if ("(Intercept)" %in% rownames(coef_storage)) coef_b["(Intercept)"] else 0
  betas <- coef_b[rownames(coef_storage) != "(Intercept)"]

  common_vars <- intersect(names(betas), colnames(X_test))
  betas <- betas[common_vars]
  X_test_num <- as.matrix(X_test[, common_vars, drop = FALSE])

  lp <- drop(X_test_num %*% betas + intercept)
  probs <- 1 / (1 + exp(-lp))

  y <- y_test   # fixed test labels

  thresh_seq <- seq(0.01, 0.99, by = 0.01)
  f1_scores <- sapply(thresh_seq, function(t) {
    preds_t <- ifelse(probs > t, 1, 0)
    safe_f1(preds_t, y)
  })

  thresholds$F1[b] <- if (all(is.na(f1_scores))) NA else thresh_seq[which.max(f1_scores)]
}

summary(na.omit(thresholds$F1))

saveRDS(thresholds, file = "boot-grLASSO-threshold-strat1.rds")
```

```{r boot grLASSO threshold READ}
thresholds <- readRDS("boot-grLASSO-threshold-strat1.rds")

summary(thresholds$F1)
median(thresholds$F1, na.rm = TRUE)
mean(thresholds$F1, na.rm = TRUE)
sd(thresholds$F1, na.rm = TRUE)

median_thresh_grLASSO <- median(thresholds$F1, na.rm = TRUE)
```

```{r grLASSO evaluation metrics}
X_test <- as.matrix(X_test)

evaluate_model <- function(model, X, y, f1_thresh = 0.99) {
  
  X <- as.matrix(X)
  
  probs <- predict(model, X,
                   type   = "response",
                   lambda = model$lambda.min)
  probs <- as.numeric(probs)
  
  preds <- ifelse(probs > f1_thresh, 1, 0)
  cm <- table(Predicted = preds, Actual = y)
  
  TP <- cm["1","1"]; TN <- cm["0","0"]
  FP <- cm["1","0"]; FN <- cm["0","1"]
  
  accuracy    <- (TP + TN) / sum(cm)
  sensitivity <- TP / (TP + FN)
  specificity <- TN / (TN + FP)
  precision   <- TP / (TP + FP)
  f1          <- 2 * precision * sensitivity / (precision + sensitivity)
  
  brier   <- mean((probs - y)^2)
  logloss <- -mean(y * log(probs) + (1 - y) * log(1 - probs))
  
  roc_obj <- pROC::roc(y, probs)
  auc_val <- pROC::auc(roc_obj)
  
  pr_obj <- PRROC::pr.curve(
  scores.class0 = probs[y == 1],
  scores.class1 = probs[y == 0],
  curve = FALSE
  )
  pr_auc <- pr_obj$auc.integral

  return(list(Accuracy = accuracy,
              Sensitivity = sensitivity,
              Specificity = specificity,
              Precision = precision,
              F1 = f1,
              PRAUC = pr_auc,
              AUC         = auc_val,
              Brier       = brier,
              LogLoss     = logloss,
              F1Thresh    = f1_thresh
            ))
}


metricsgroupLASSO <- evaluate_model(cvfit_grLASSO_reduced, X_test_grLASSO_reduced, y_test, f1_thresh = median_thresh_grLASSO)
metricsgroupLASSO
```


### 1G.3 Proposed Model: grSCAD

```{r reduced grSCAD model}
X_train_grSCAD_reduced <- model.matrix(Landslide ~ Slope + PGAComp +  Lithology + SoilType,
                     data = train_data)[, -1]  
X_test_grSCAD_reduced <- model.matrix(Landslide ~ Slope + PGAComp + Lithology + SoilType,
                                       data = test_data)[, -1]

cols <- colnames(X_train_grSCAD_reduced)

group_grSCAD_reduced <- integer(length(cols))

group_grSCAD_reduced[grep("^Slope$", cols)]    <- 1
group_grSCAD_reduced[grep("^PGAComp$", cols)]  <- 2
group_grSCAD_reduced[grep("^Lithology", cols)] <- 3
group_grSCAD_reduced[grep("^SoilType", cols)]  <- 4

table(group_grSCAD_reduced)
cbind(cols, group_grSCAD_reduced) 

fit_grSCAD_reduced <- grpreg(
  X_train_grSCAD_reduced, y_train,
  group = group_grSCAD_reduced,
  penalty = "grLasso",
  family = "binomial"
)

set.seed(123)
cvfit_grSCAD_reduced <- cv.grpreg(
  X_train_grSCAD_reduced, y_train,
  group = group_grSCAD_reduced,
  penalty = "grLasso",  
  family = "binomial",
  seed = 123
)

lambda_min_grSCAD_reduced <- cvfit_grSCAD_reduced$lambda.min
cvfit_grSCAD_reduced |> plot()
lambda_min_grSCAD_reduced


coef_min_grSCAD_reduced <- coef(cvfit_grSCAD_reduced, lambda = lambda_min_grSCAD_reduced)
coef_min_grSCAD_reduced    
```

```{r boot grLASSO threshold CODE}
results <- readRDS("boot-grSCAD-results-strat1.rds")
coef_storage    <- results$coefficients
prediction_list <- results$predictions
actual_list     <- results$actuals

inclusion_freq <- rowMeans(coef_storage != 0) * 100

selected_levels <- names(inclusion_freq[inclusion_freq >= 80])

X_train_reduced <- as.matrix(X_train[, intersect(colnames(X_train), selected_levels), drop = FALSE])

cv_scad_reduced <- cv.glmnet(
  X_train_reduced, y_train,
  family = "binomial", alpha = 1,
  standardize = FALSE
)

X_test <- as.data.frame(X_test)  # ensure data frame
missing_cols <- setdiff(colnames(X_train_reduced), colnames(X_test))
for (mc in missing_cols) {
  X_test[[mc]] <- 0
}
X_test_reduced <- as.matrix(X_test[, colnames(X_train_reduced), drop = FALSE])

stopifnot(nrow(X_test_reduced) == length(y_test))



thresholds <- data.frame(Rep = 1:length(prediction_list),
                         F1 = NA_real_)

for (i in seq_along(prediction_list)) {
  probs <- as.numeric(prediction_list[[i]])
  y     <- actual_list[[i]]

  thresh_seq <- seq(0.01, 0.99, by = 0.01)
  f1_scores <- sapply(thresh_seq, function(t) {
    preds_t <- ifelse(probs > t, 1, 0)

    cm_t <- table(
      Predicted = factor(preds_t, levels = c(0,1)),
      Actual    = factor(y, levels = c(0,1))
    )

    TP_t <- cm_t["1","1"]; FN_t <- cm_t["0","1"]; FP_t <- cm_t["1","0"]
    prec_t <- ifelse(TP_t + FP_t == 0, 0, TP_t / (TP_t + FP_t))
    rec_t  <- ifelse(TP_t + FN_t == 0, 0, TP_t / (TP_t + FN_t))
    if (prec_t + rec_t == 0) return(0)
    2 * prec_t * rec_t / (prec_t + rec_t)
  })

  thresholds$F1[i] <- thresh_seq[which.max(f1_scores)]
}


saveRDS(thresholds, file = "boot-grSCAD-threshold-strat1.rds")
```


```{r boot grLASSO threshold READ}
thresholds <- readRDS("boot-grSCAD-threshold-strat1.rds")

summary(thresholds$F1)
median(thresholds$F1, na.rm = TRUE)
mean(thresholds$F1, na.rm = TRUE)
sd(thresholds$F1, na.rm = TRUE)

median_thresh_grSCAD <- median(thresholds$F1, na.rm = TRUE)
```


```{r grSCAD evaluation metrics}
X_test <- as.matrix(X_test)

evaluate_model <- function(model, X, y, f1_thresh = 0.51) {
  X <- as.matrix(X)
  
  probs <- predict(model, X,
                   type   = "response",
                   lambda = model$lambda.min)
  probs <- as.numeric(probs)
  
  preds <- ifelse(probs > f1_thresh, 1, 0)
  cm <- table(Predicted = preds, Actual = y)
  
  TP <- cm["1","1"]; TN <- cm["0","0"]
  FP <- cm["1","0"]; FN <- cm["0","1"]
  
  accuracy    <- (TP + TN) / sum(cm)
  sensitivity <- TP / (TP + FN)
  specificity <- TN / (TN + FP)
  precision   <- TP / (TP + FP)
  f1          <- 2 * precision * sensitivity / (precision + sensitivity)
  
  brier   <- mean((probs - y)^2)
  logloss <- -mean(y * log(probs) + (1 - y) * log(1 - probs))
  
  roc_obj <- pROC::roc(y, probs)
  auc_val <- pROC::auc(roc_obj)
  
  pr_obj <- PRROC::pr.curve(
  scores.class0 = probs[y == 1],
  scores.class1 = probs[y == 0],
  curve = FALSE
  )
  pr_auc <- pr_obj$auc.integral

  return(list(Accuracy = accuracy,
              Sensitivity = sensitivity,
              Specificity = specificity,
              Precision = precision,
              F1 = f1,
              PRAUC = pr_auc,
              AUC         = auc_val,
              Brier       = brier,
              LogLoss     = logloss,
              F1Thresh    = f1_thresh
            ))
}


metricsgroupSCAD <- evaluate_model(cvfit_grSCAD_reduced, X_test_grSCAD_reduced, y_test, f1_thresh = median_thresh_grSCAD)
metricsgroupSCAD
```

## 1H Cross Validation

### 1H.1 Reduced LR Model's Cross-Validated Performance

```{r reduced lr CV}
set.seed(123)

cv_folds <- createFolds(data$Landslide, k = 10, list = TRUE, returnTrain = FALSE)

cv_results <- data.frame(
  Accuracy = numeric(10),
  Sensitivity = numeric(10),
  Specificity = numeric(10),
  Precision = numeric(10),
  F1 = numeric(10),
  AUC = numeric(10),
  PR_AUC = numeric(10),
  Brier = numeric(10),
  LogLoss = numeric(10)
)

coef_list <- vector("list", 10)


for (i in 1:10) {
  test_idx <- cv_folds[[i]]
  test_fold <- data[test_idx, ]
  train_fold <- data[-test_idx, ]
  
  # --- Set reference levels for categorical predictors (must exist in levels) ---
  train_fold$SoilType   <- relevel((train_fold$SoilType), ref = "Loam")
  train_fold$Lithology  <- relevel((train_fold$Lithology), ref = "Recent deposits")
  train_fold$Aspect     <- relevel((train_fold$Aspect), ref = "North")
  
  test_fold$SoilType    <- relevel((test_fold$SoilType), ref = "Loam")
  test_fold$Lithology   <- relevel((test_fold$Lithology), ref = "Recent deposits")
  test_fold$Aspect      <- relevel((test_fold$Aspect), ref = "North")
  
  # Fit LR on training folds
  fit <- glm(Landslide ~ SoilType + Slope + Lithology + PGAComp + Elev,
             data = train_fold, family = "binomial")
  
  # Save coefficients for this fold
  coef_list[[i]] <- coef(fit)
  
  # Predict on held-out fold
  preds <- predict(fit, newdata = test_fold, type = "response")
  pred_class <- ifelse(preds > 0.56, 1, 0)
  
  # Confusion matrix
  cm <- table(Predicted = pred_class, Actual = test_fold$Landslide)
  
  TP <- cm["1","1"]; TN <- cm["0","0"]
  FP <- cm["1","0"]; FN <- cm["0","1"]
  
  acc <- (TP + TN) / sum(cm)
  sens <- TP / (TP + FN)
  spec <- TN / (TN + FP)
  prec <- TP / (TP + FP)
  f1 <- 2 * prec * sens / (prec + sens)
  
  # ROC & AUC
  roc_obj <- roc(test_fold$Landslide, preds)
  auc_val <- auc(roc_obj)
  
  pr_obj <- PRROC::pr.curve(
  scores.class0 = preds[test_fold$Landslide == 1],
  scores.class1 = preds[test_fold$Landslide == 0],
  curve = FALSE
  )
  pr_auc <- pr_obj$auc.integral
  
  
  # Brier score
  brier <- mean((preds - test_fold$Landslide)^2)
  
  # LogLoss
  eps <- 1e-15
  logloss <- -mean(test_fold$Landslide * log(pmax(preds, eps)) +
                   (1 - test_fold$Landslide) * log(pmax(1 - preds, eps)))
  
  # Store metrics
  #cv_results[i,] <- c(acc, sens, spec, prec, f1, auc_val, brier, logloss)
  cv_results[i,] <- c(acc, sens, spec, prec, f1, auc_val, pr_auc, brier, logloss)
}

# Average metrics across folds
cv_summary_lr <- colMeans(cv_results, na.rm = TRUE)

# Combine coefficients across folds
coef_mat <- do.call(cbind, coef_list)

coef_summary_cv_lr <- data.frame(
  Term   = rownames(coef_mat),
  Mean   = rowMeans(coef_mat, na.rm = TRUE),
  Median = apply(coef_mat, 1, median, na.rm = TRUE),
  SD     = apply(coef_mat, 1, sd, na.rm = TRUE)
)

cv_summary_lr
coef_summary_cv_lr
```


### 1H.2 Reduced grLASSO Model's Cross-Validated Performance

```{r reduced grLASSO CV}
set.seed(123)

data$SoilType   <- relevel(factor(data$SoilType), ref = "Loam")
data$Lithology  <- relevel(factor(data$Lithology), ref = "Recent deposits")
data$Aspect     <- relevel(factor(data$Aspect), ref = "North")

X <- model.matrix(~ Elev + Slope + PGAComp + Aspect + Lithology + SoilType, 
                  data = data)[, -1]
y <- data$Landslide

cols <- colnames(X)
group <- integer(length(cols))

group[grep("^Elev$", cols)]                <- 1
group[grep("^Slope$", cols)]               <- 2
group[grep("^PGAComp$", cols)]             <- 3
group[grep("^Aspect", cols)]     <- 4
group[grep("^Lithology", cols)]  <- 5
group[grep("^SoilType", cols)]   <- 6

print(table(group))
print(cbind(cols, group))


# 10-fold CV
cv_folds <- createFolds(y, k = 10, list = TRUE)

cv_results <- data.frame(
  Accuracy = numeric(10),
  Sensitivity = numeric(10),
  Specificity = numeric(10),
  Precision = numeric(10),
  F1 = numeric(10),
  AUC = numeric(10),
  PR_AUC = numeric(10),
  Brier = numeric(10),
  LogLoss = numeric(10)
)

coef_list <- vector("list", 10)

for (i in 1:10) {
  test_idx  <- cv_folds[[i]]
  train_idx <- setdiff(seq_len(nrow(data)), test_idx)
  
  X_train <- X[train_idx, ]
  y_train <- y[train_idx]
  
  X_test  <- X[test_idx, ]
  y_test  <- y[test_idx]
  
  
  cvfit <- cv.grpreg(X_train, y_train,
                     group = group,
                     family = "binomial",
                     penalty = "grLasso")
  
  lambda_min <- cvfit$lambda.min
  fit <- cvfit$fit
  
  coef_list[[i]] <- coef(fit, lambda = lambda_min)
  
  preds <- predict(fit, X_test, type = "response", lambda = lambda_min)
  eps <- 1e-15
  preds <- pmin(pmax(preds, eps), 1 - eps)

  pred_class <- ifelse(preds > 0.99, 1, 0)
  cm <- table(Predicted = pred_class, Actual = y_test)
  
  TP <- cm["1","1"]; TN <- cm["0","0"]
  FP <- cm["1","0"]; FN <- cm["0","1"]
  
  acc  <- (TP + TN) / sum(cm)
  sens <- TP / (TP + FN)
  spec <- TN / (TN + FP)
  prec <- TP / (TP + FP)
  f1   <- 2 * prec * sens / (prec + sens)
  
  roc_obj <- roc(y_test, preds)
  auc_val <- auc(roc_obj)
  
  pr_obj <- PRROC::pr.curve(
  scores.class0 = preds[y_test == 1],
  scores.class1 = preds[y_test == 0],
  curve = FALSE
  )
  pr_auc <- pr_obj$auc.integral
  
  
  brier <- mean((preds - y_test)^2)
  logloss <- -mean(y_test * log(preds) + (1 - y_test) * log(1 - preds))
  
  cv_results[i,] <- c(acc, sens, spec, prec, f1, auc_val, pr_auc, brier, logloss)
}

cv_summary_grLasso <- colMeans(cv_results, na.rm = TRUE)

coef_mat <- do.call(cbind, coef_list)
coef_summary_cv_grLasso <- data.frame(
  Term   = rownames(coef_mat),
  Mean   = rowMeans(coef_mat, na.rm = TRUE),
  Median = apply(coef_mat, 1, median, na.rm = TRUE),
  SD     = apply(coef_mat, 1, sd, na.rm = TRUE)
)


cv_summary_grLasso
coef_summary_cv_grLasso 
```

### 1H.3 Reduced grSCAD Model's Cross-Validated Performance

```{r reduced grSCAD CV}
set.seed(123)

data$SoilType   <- relevel(factor(data$SoilType), ref = "Loam")
data$Lithology  <- relevel(factor(data$Lithology), ref = "Recent deposits")
data$Aspect     <- relevel(factor(data$Aspect), ref = "North")

X <- model.matrix(~ SoilType + Slope + Lithology + PGAComp, data = data)[, -1]
y <- data$Landslide

cols <- colnames(X)
group <- integer(length(cols))
group[grep("^Slope$", cols)]     <- 1
group[grep("^PGAComp$", cols)]   <- 2
group[grep("^Lithology", cols)]  <- 3
group[grep("^SoilType", cols)]   <- 4

# 10-fold CV
cv_folds <- createFolds(y, k = 10, list = TRUE)

cv_results <- data.frame(
  Accuracy = numeric(10),
  Sensitivity = numeric(10),
  Specificity = numeric(10),
  Precision = numeric(10),
  F1 = numeric(10),
  AUC = numeric(10),
  PR_AUC = numeric(10),
  Brier = numeric(10),
  LogLoss = numeric(10)
)

coef_list <- vector("list", 10)

for (i in 1:10) {
  test_idx  <- cv_folds[[i]]
  train_idx <- setdiff(seq_len(nrow(data)), test_idx)
  
  X_train <- X[train_idx, ]
  y_train <- y[train_idx]
  
  X_test  <- X[test_idx, ]
  y_test  <- y[test_idx]
  
  cvfit <- cv.grpreg(X_train, y_train,
                     group = group,
                     family = "binomial",
                     penalty = "grSCAD")
  
  lambda_min <- cvfit$lambda.min
  fit <- cvfit$fit
  
  coef_list[[i]] <- coef(fit, lambda = lambda_min)
  
  preds <- predict(fit, X_test, type = "response", lambda = lambda_min)
  eps <- 1e-15
  preds <- pmin(pmax(preds, eps), 1 - eps)

  pred_class <- ifelse(preds > 0.51, 1, 0)
  cm <- table(Predicted = pred_class, Actual = y_test)
  
  TP <- cm["1","1"]; TN <- cm["0","0"]
  FP <- cm["1","0"]; FN <- cm["0","1"]
  
  acc  <- (TP + TN) / sum(cm)
  sens <- TP / (TP + FN)
  spec <- TN / (TN + FP)
  prec <- TP / (TP + FP)
  f1   <- 2 * prec * sens / (prec + sens)
  
  roc_obj <- roc(y_test, preds)
  auc_val <- auc(roc_obj)
  
  pr_obj <- PRROC::pr.curve(
  scores.class0 = preds[y_test == 1],
  scores.class1 = preds[y_test == 0],
  curve = FALSE
  )
  pr_auc <- pr_obj$auc.integral
  
  brier <- mean((preds - y_test)^2)
  logloss <- -mean(y_test * log(preds) + (1 - y_test) * log(1 - preds))
  
  cv_results[i,] <- c(acc, sens, spec, prec, f1, auc_val, pr_auc, brier, logloss)
}

cv_summary_grSCAD <- colMeans(cv_results, na.rm = TRUE)

coef_mat <- do.call(cbind, coef_list)
coef_summary_cv_grSCAD <- data.frame(
  Term   = rownames(coef_mat),
  Mean   = rowMeans(coef_mat, na.rm = TRUE),
  Median = apply(coef_mat, 1, median, na.rm = TRUE),
  SD     = apply(coef_mat, 1, sd, na.rm = TRUE)
)


cv_summary_grSCAD
coef_summary_cv_grSCAD 
```

## 2I Mapping

### 2I.1 Map via LR model 
```{r mapping and classification using reduced LR model}
# standard logistic regression
lr_model <- glm(Landslide ~ Elev + Slope + SoilType + Lithology + PGAComp,
                data = data,
                family = binomial)
data$Prob <- predict(lr_model, newdata = data, type = "response")

nc_boundary <- st_read("north_cotabato_gadm.shp")
pts <- st_as_sf(data, coords = c("Long", "Lat"), crs = 4326)
pts_joined <- st_join(pts, nc_boundary)
pts_in_nc <- pts_joined[!is.na(pts_joined$NAME_1), ]   
data_sp <- as(pts_in_nc, "Spatial")

grd_sf <- st_make_grid(nc_boundary, cellsize = 0.001, what = "centers") # 0.001 is 100mX100m
grd_sf <- st_intersection(grd_sf, nc_boundary)
grd <- as(grd_sf, "Spatial")
sp::gridded(grd) <- TRUE

idw_result <- gstat::idw(Prob ~ 1, data_sp, newdata = grd)
idw_raster <- raster(idw_result)


# Classification with Jenks Natural Breaks 
vals <- values(idw_raster)
vals <- vals[!is.na(vals)]

jenks <- classIntervals(vals, n = 5, style = "jenks")
breaks <- jenks$brks

classified <- cut(idw_raster,
                  breaks = breaks,
                  include.lowest = TRUE)

class_labels <- c("Very Low", "Low", "Moderate", "High", "Very High")
levels(classified) <- data.frame(ID = 1:5, Class = class_labels)

my_colors1 <- c("#2DC937", "#99C140", "#E7B416", "#DB7B2B", "#CC3232")

png("reduced_lr_LSM_data1.png", width = 1200, height = 900, res = 150)

plot(classified,
     col = my_colors1,
     legend = FALSE,
     main = "Landslide Susceptibility Map (Reduced LR)")

plot(st_geometry(nc_boundary),
     add = TRUE,
     border = "black",
     lwd = 1)

legend("bottomleft",
       legend = class_labels,
       fill = my_colors1,
       title = "Susceptibility",
       bty = "o")

dev.off()
```

### 2I.2 Map via grSCAD CV model
```{r mapping and classification using reduced grSCAD CV model}
X_grSCAD_reduced_lsm <- model.matrix(Landslide ~ Slope + PGAComp + Lithology + SoilType,
                                     data = data)[, -1]

cols <- colnames(X_grSCAD_reduced_lsm)

group_grSCAD_reduced_lsm <- integer(length(cols))
group_grSCAD_reduced_lsm[grep("^Slope$", cols)]    <- 1
group_grSCAD_reduced_lsm[grep("^PGAComp$", cols)]  <- 2
group_grSCAD_reduced_lsm[grep("^Lithology", cols)] <- 3
group_grSCAD_reduced_lsm[grep("^SoilType", cols)]  <- 4

# Group SCAD logistic regression with CV
set.seed(123)
cvfit_grSCAD_reduced_lsm <- cv.grpreg(
  X_grSCAD_reduced_lsm, data$Landslide,
  group   = group_grSCAD_reduced_lsm,
  penalty = "grSCAD",
  family  = "binomial",
  seed    = 123
)

data$Prob <- predict(cvfit_grSCAD_reduced_lsm,
                     X_grSCAD_reduced_lsm,
                     type = "response",
                     lambda = cvfit_grSCAD_reduced_lsm$lambda.min)

nc_boundary <- st_read("north_cotabato_gadm.shp")
pts <- st_as_sf(data, coords = c("Long", "Lat"), crs = 4326)
pts_joined <- st_join(pts, nc_boundary)
pts_in_nc <- pts_joined[!is.na(pts_joined$NAME_1), ]
data_sp <- as(pts_in_nc, "Spatial")

grd_sf <- st_make_grid(nc_boundary, cellsize = 0.001, what = "centers") # 0.001 is 100mX100m
grd_sf <- st_intersection(grd_sf, nc_boundary)

grd <- as(grd_sf, "Spatial")
sp::gridded(grd) <- TRUE

idw_result <- gstat::idw(Prob ~ 1, data_sp, newdata = grd)
idw_raster <- raster(idw_result)


# Classification with Jenks Natural Breaks
vals <- values(idw_raster)
vals <- vals[!is.na(vals)]

jenks <- classIntervals(vals, n = 5, style = "jenks")
breaks <- jenks$brks

classified <- cut(idw_raster,
                  breaks = breaks,
                  include.lowest = TRUE)

class_labels <- c("Very Low", "Low", "Moderate", "High", "Very High")
levels(classified) <- data.frame(ID = 1:5, Class = class_labels)

my_colors1 <- c("#2DC937", "#99C140", "#E7B416", "#DB7B2B", "#CC3232")

png("reduced_grSCAD_LSM_data1.png", width = 1200, height = 900, res = 150)

plot(classified,
     col = my_colors1,
     legend = FALSE,
     main = "Landslide Susceptibility Map (Reduced SCAD)")

plot(st_geometry(nc_boundary),
     add = TRUE,
     border = "black",
     lwd = 1)

legend("bottomleft",
       legend = class_labels,
       fill = my_colors1,
       title = "Susceptibility",
       bty = "o")

dev.off()
```
